Four-dimensional emittance measurement at the Spallation Neutron Source

### 2022-07-24 main.tex
Final edits

%% For including figures, graphicx.sty has been loaded in
%% elsarticle.cls. If you prefer to use the old commands
%% please give \usepackage{epsfig}
%% The amssymb package provides various useful mathematical symbols
\usepackage{amssymb}
%% The amsthm package provides extended theorem environments
%% \usepackage{amsthm}
\usepackage{subcaption} % subfigures
\usepackage{gensymb} % \degree symbol
\usepackage{bm}
\usepackage{amsmath}
\usepackage{dblfloatfix} % Put figures at bottom of page
\usepackage{booktabs} % For tables
\usepackage{lineno} % Line numbers
% \linenumbers

@@

\title{Fixed-optics four-dimensional emittance measurement at the Spallation Neutron Source}

@@

\begin{abstract}
A coasting hadron beam with an elliptical transverse profile, uniform charge density, and small transverse four-dimensional (4D) emittance could improve accelerator performance in several contexts. A phase space painting method to generate such a distribution is being tested in the Spallation Neutron Source (SNS) accumulator ring. A critical component of these efforts is to measure the 4D emittance of the beam. The 4D emittance can be reconstructed from measured beam profiles in two ways: in the multi-optics method, the optics between a reconstruction and measurement location are varied; in the fixed-optics method, multiple measurement locations are used without modifying the optics. The fixed-optics is much faster but can potentially lead to unacceptable bias and uncertainty in the reconstructed 4D emittance. In this paper, we discuss the implementation of a variant of the multi-optics method using the four available wire-scanners near the SNS target, as well as the modification of the wire-scanner region to utilize the fixed-optics method. We then demonstrate the usefulness of the fixed-optics method by reconstructing the 4D emittance evolution during accumulation in the SNS ring.
\end{abstract}

@@

The available diagnostics in the SNS ring-target beam transport line (RTBT) are four wire-scanners that measure 1D projections of the distribution onto a horizontal, vertical, and diagonal axis in the transverse plane. The covariance matrix, and hence the root-mean-square emittances, can be reconstructed from at least four wire-scanner measurements with different transfer matrices connecting the measurement and reconstruction locations \cite{book:Minty2003}.\footnote{Reconstruction of the transverse phase space distribution is possible using 2D projections of the distribution on a screen \cite{Hock2013, Wang2019, Wolski2020}. This may be possible using the SNS target imaging system \cite{Blokland2010} but is not pursued here.} In general, the transfer matrix elements can be controlled by varying the machine optics (multi-optics method), the measurement location (fixed-optics method), or any combination of the two \cite{Prat2014}. With four available wire-scanners, the fixed-optics method is possible and preferred in our case due to its speed; however, previous studies have shown that the method can be sensitive to errors \cite{Raimondi1993, Woodley2000, Agapov2007, Faus-Golfe2016}.

The purpose of this work is to determine the limitations of the fixed-optics method in the SNS. The reconstruction method is reviewed in Section \ref{sec:Four-dimensional emittance measurement}; the implementation of the multi-optics and fixed-optics methods in the SNS is discussed in Section \ref{sec:Implementation in the SNS}; the fixed-optics method is demonstrated in Section \ref{sec:Example application} by measuring the 4D emittance evolution during accumulation in the SNS ring.

@@

The SNS produces short neutron pulses by colliding a proton beam with a liquid mercury target. The beam starts as a continuous wave of H$^-$ ions which is chopped and then bunched in a 402.5 MHz radio-frequency quadrupole (RFQ), forming microsecond-long minipulses. Each minipulse is accelerated in a linac to a kinetic energy of 1 GeV, stripped of its electrons using a carbon foil, and injected into a ring. Approximately one thousand minipulses are accumulated in the ring over $10^{-3}$ seconds. Finally, the proton beam is extracted and guided through the ring-target beam transport line (RTBT) to the target. This process repeats sixty times per second, resulting in an average beam power of 1.5 megawatts. A comprehensive description of the SNS is given in \cite{Henderson2014}; several design parameters of the SNS ring and RTBT are listed in Table~\ref{tab:SNS}; an overhead view of the ring and RTBT is shown in Fig.~\ref{fig:1}.

%
\begin{table}[!b]
    \centering
    \begin{tabular}{ll}
        \toprule
        {Quantity} & {Value} \\
        \midrule
        Ring length & 248 [m] \\
        RTBT length & 151 [m] \\
        Beam kinetic energy & 1 [GeV] \\
        Average beam power & 1.4 [MW] \\
        Repetition rate & 60 [Hz] \\
        Transverse tunes ($\nu_x$, $\nu_y$) & (6.23, 6.20) \\
        Longitudinal tune & $7 \times 10^{-4}$ \\
        \midrule 
    \end{tabular}
    \caption{Several design parameters of the SNS ring and RTBT.}
    \label{tab:SNS}
\end{table}
%

Diagnostics in the SNS ring are limited to beam position and loss monitors. Beam profile measurements are performed in the RTBT using four wire-scanners, labeled WS20, WS21, WS23, and WS24, the locations of which are shown in Fig.~\ref{fig:1}. Each wire-scanner consists of three 100$\mu m$-thick tungsten wires mounted on a fork at forty-five degree angles relative to each other. The intensity of secondary electron emission from each wire is recorded as the fork moves across the beam, thus measuring the 1D projection of the beam distribution onto the $x$, $y$ and $u$ axes, where $u$ is tilted 45 degrees above the $x$ axis. The second-order moments $\langle{xx}\rangle$, $\langle{yy}\rangle$, and $\langle{uu}\rangle$ are estimated from the profiles by averaging over the wire positions weighted by the signal amplitude. The cross-plane moment $\langle{xy}\rangle$ is thus measured indirectly: if the tilt angle of the $u$ axis is $\phi$, then $u = x\cos\phi + y\sin\phi$ and 
%
\begin{equation}
    \langle{xy}\rangle = \frac{\langle{uu}\rangle - \langle{xx}\rangle \cos^2\phi - \langle{yy}\rangle \sin^2\phi}{2\sin\phi\cos\phi}.
\end{equation}
%
The wire-scanners are run in parallel, taking approximately five minutes to move across the beam and return to their original positions. Their step size is 3 mm and their dynamic range (the ratio of maximum to minimum measured signal) is approximately $100$.\footnote{The root-mean-square beam sizes may depend on beam tails/halo that cannot be measured by the RTBT wire-scanners. Investigation into the effect of beam halo on emittance measurements in the RTBT is left as future work.} They are run at a beam pulse frequency of 1 Hz.\footnote{Each data point corresponds to a separate beam pulse, so the measurement relies on pulse-to-pulse stability.} 
%
\begin{figure*}[!t]
    \centering
    \begin{subfigure}{0.375\columnwidth}
        % \raisebox{20pt}{\includegraphics[width=\textwidth]{fig1a.pdf}}
        \includegraphics[width=\textwidth]{fig1a.pdf}
    \end{subfigure}
    % \hspace{1cm}}
    \begin{subfigure}{0.61\columnwidth}
        \includegraphics[width=\textwidth]{fig1b.pdf}
    \end{subfigure}
    \caption{Left: The accumulator ring and ring-target-beam-transport (RTBT) sections of the Spallation Neutron Source (SNS). Right: $\beta$ functions (top), phase advances (middle), and focusing element and wire-scanner positions (bottom) in the second half of the RTBT. The red box outlines the wire-scanner region.}
    \label{fig:1}
\end{figure*}
%

@@

We first implemented a variant of the multi-optics method in the SNS. There is considerably more freedom in the choice of optics used in the 4D reconstruction than in 2D reconstruction. We do not consider the general case of coupled optics in this work since the RTBT optics are uncoupled. Thus, the problem reduces to sampling the space of horizontal/vertical phase advances from the reconstruction location to the measurement location such that the measurement error is minimized.\footnote{
    In 2D reconstructions, the phase advance from the reconstruction location to the measurement location is typically varied uniformly within a $180\degree$ range. The error is expected to be minimized when the projection angles are spread uniformly between 0 and $\pi$ (the projection angles are only equal to the phase advances in normalized phase space, but it is often convenient to vary the phase advances instead) \cite{Hock2011}. A clear geometric argument is based on the fact that the measured beam sizes define a set of lines that bound an ellipse in 2D phase space at the reconstruction location, so the reconstruction error will be minimized if these lines are evenly distributed around the ellipse. But this argument is less clear in a 4D reconstruction: the measurements of $\left\{\langle{xx}\rangle, \langle{yy}\rangle, \langle{xy}\rangle\right\}$ define a set of surfaces that bound an ellipsoid in 4D phase space at the reconstruction location, and it is not entirely clear how to "evenly distribute" these surfaces around the ellipsoid.
} We chose to follow Prat and Aiba \cite{Prat2014}: in the first half of the scan, the horizontal phase advance is varied while the vertical phase advance is held fixed; in the second half of the scan, the vertical phase advance is varied while the horizontal phase advance is held fixed. Since the four SNS wire-scanners are already spaced somewhat evenly in phase advance, maximal phase coverage is possible by varying the phase advances at each wire-scanner in a $30\degree$ window; however, control of the phase advances between each wire-scanner is limited due to the shared power supplies of the RTBT quadrupoles. There are two power supplies in the wire-scanner region — one controls \{QH18, QH20, QH22, QH24\} and the other controls \{QV19, QV21, QV23, QV25\} — while the last five quadrupoles are controlled independently. In addition to these constraints, the $\beta$ functions must be kept small in the wire-scanner region and must remain close to their nominal values at the target.

@@

This multi-optics method was tested on a fully accumulated production beam in the SNS using ten measurements (forty profiles), which were collected in one hour. The reconstructed Twiss parameters, emittances, and best-fit ellipses in the $x$-$x'$ and $y$-$y'$ planes (normalized by the reconstructed Twiss parameters) are displayed in Fig.~\ref{fig:2}.
%
\begin{figure}[b!]
    \centering
    \includegraphics[width=0.65\columnwidth]{fig2.pdf}
    % \par\medskip
    \begin{tabular}{lll}
        \small\textbf{Parameter} & \small\textbf{Measurement} & \small\textbf{Model} \\
        % \midrule
        \small$\beta_x$ [m/rad] & \small22.06 $\pm$ 0.29 & \small22.00 \\
        \small$\beta_y$ [m/rad] & \small4.01 $\pm$ 0.02 & \small3.81 \\
        \small$\alpha_x$ & \small2.33 $\pm$ 0.04 & \small2.37 \\
        \small$\alpha_y$ & \small-0.49 $\pm$ 0.01 & \small-0.60 \\
        \small$\varepsilon_1$ [mm~mrad] & \small33.02 $\pm$ \small0.05 & - \\
        \small$\varepsilon_2$ [mm~mrad] & \small25.67 $\pm$ \small1.03 & - \\
        \small$\varepsilon_x$ [mm~mrad] & \small32.85 $\pm$ \small0.05 & - \\
        \small$\varepsilon_y$ [mm~mrad] & \small25.87 $\pm$ \small0.12 & - \\
      \end{tabular}
    % \par\medskip
    \caption{Reconstructed beam parameters and graphical output (in normalized phase space) from a multi-optics emittance measurement of a production beam.}
    \label{fig:2}
\end{figure}
%
The uncertainties in the parameters were calculated by propagating the standard deviations of the ten reconstructed moments obtained from the LLSQ estimator \cite{Faus-Golfe2016}. The reconstructed Twiss parameters are close to the model parameters computed from the linear transfer matrices of the ring and RTBT, showing that the beam was matched to the nominal optics when it was extracted from the ring. The intrinsic emittances are almost equal to the apparent emittances, showing that there was very little cross-plane correlation in the beam. This is expected for a production beam.

A comprehensive study of errors in the multi-optics 4D emittance measurement was completed at the SwissFEL Injector Test Facility (SITF) by Prat and Aiba in \cite{Prat2014}. They considered errors in the measured moments, quadrupole field and alignment errors, beam energy errors, beam mismatch at the reconstruction point, and dispersion/chromaticity \cite{Mostacci2012}, concluding that the method remained accurate and reporting $< 5\%$ uncertainty in the intrinsic emittances. We initially performed similar studies using envelope tracking to estimate the reconstruction errors in the RTBT \cite{Hoover2021-IPAC}. We considered uncertainty in the beam moments ($\pm 5\%$) quadrupole tilt angles ($\pm\%$ 1 mrad), quadrupole field strengths ($\pm$ 1\%), beam kinetic energy ($\pm$ 3 MeV for a 1 GeV beam), and beam Twiss parameters at the reconstruction point ($\pm$ 5\%), with all other effects turned off. The emittances were successfully reconstructed with systematic and random errors of a few percent. Most of these uncertainties had only a small affect on the reconstructed intrinsic emittances, with the largest coming from beam mismatch and errors in the measured beam size. Therefore, we only consider these sources of uncertainty going forward.\footnote{Space charge forces, which can render the method invalid for high-perveance beams \cite{Anderson2002}, can be neglected since the space charge tune shift in the ring is around 3\% and the distance from the reconstruction location to the measurement location is much less than the length of the ring.} From repeated wire-scanner measurements with the same machine setup (see Fig.~\ref{fig:3}), we estimated the error in the measured moments to be only 3\%.
%
\begin{figure}
    \centering
    \includegraphics[width=0.5\textwidth]{profiles.pdf}
    \caption{Measured profiles of a production beam at WS20 in the RTBT.}
    \label{fig:3}
\end{figure}
%

@@

Th downside of the multi-optics method is its long execution time, for the following reasons. First, we are not only interested in the beam emittances at a single time but are also interested in the emittance evolution throughout accumulation, which conveys valuable information about the beam dynamics and allows qualitative comparison with computer simulation. Second, it would be beneficial to quickly evaluate various machine states: although computer simulation can be used as a guide in choosing the experimental parameters, online tuning is expected to be necessary since small changes to the accelerator may have large effects on the painted phase space distribution. Third, the time reserved for accelerator physics experiments at the SNS is limited, and our initial experiments will require substantial setup time due to modification of the beam energy. Thus, the fixed-optics method is preferred. A modest reduction in accuracy for the increase in speed is warranted since weak cross-plane correlations ($\varepsilon_1 \varepsilon_2 \approx \varepsilon_x \varepsilon_y$) are uninteresting for our purposes and do not need to be resolved.

@@

To investigate the sensitivity of the fixed-optics method, we generated a covariance matrix with $\varepsilon_1 = \varepsilon_x$ = 32 mm~mrad and $\varepsilon_2 = \varepsilon_y$ = 25 mm~mrad, close to the measured values, and with Twiss parameters matched to the lattice. This covariance matrix was tracked to the wire-scanners using the transfer matrices from the fifth step in the scan of Fig.~\ref{fig:2}, and the reconstruction was performed many times with 3\% random noise added to the $\langle{xx}\rangle$, $\langle{yy}\rangle$, and $\langle{uu}\rangle$ moments. (We ignored any uncertainty in the transfer matrix elements, which were expected to be small.) This resulted in a large fraction of failed trials, but some successful trials. Fig.~\ref{fig:4} shows the emittances in the successful trials.
%
\begin{figure}[]
    \centering
    \includegraphics[width=0.5\columnwidth]{fig3.pdf}
    \caption{Simulated emittance measurement using the nominal optics in the RTBT.}
    \label{fig:4}
\end{figure}
%
Unlike the apparent emittances, the intrinsic emittances are strongly correlated and are not centered on the correct values. Our primary concern is to reduce the latter effect, which we refer to as bias.

@@

Although most of the points are clustered near the original bias and standard deviation of 7\% and 5\%, respectively, the bias increases to nearly 15\% in some cases, which may make it difficult to resolve weak cross-plane correlation; however, the measurement should still resolve strong cross-plane correlation. This is demonstrated in the rest of the plots in Fig.~\ref{fig:6}a, in which the entire process is repeated with $\varepsilon_1 / \varepsilon_2 > 1$. The bias in the reconstruction quickly decreases — the emittances are clustered around their true values. We conclude that with modifications to the RTBT optics, the fixed-optics method should be sufficient for four-dimensional emittance measurements in the SNS.

@@

It is first necessary to describe our approach to assigning uncertainty, in the form of error bars, to the emittances reconstructed by the fixed optics method. Propagating the uncertainties in the reconstructed moments, as obtained from the LLSQ estimator, is not an option since no least-squares fit is performed — the cross-plane moments are exactly determined by the four measured profiles. Since the wire-scanners take over five minutes to measure the beam profiles, it is impossible to compute the mean and standard deviation over many trials. Our proposal is to use a similar approach as in the previous section: repeatedly perturb the measured moments in a 3\% range, each time reconstructing the covariance matrix and computing the emittances; compute the the mean and standard deviation of the emittances over all trials; plot an error bar centered at the mean, with a width equal to twice the standard deviation.

We first display the measured emittance evolution of a 1 GeV beam produced by the standard SNS painting method.\footnote{The standard painting method is often referred to as "correlated painting". The distance between the injected and circulating beam is varied as the square root of time while the angle between the beams is fixed at zero. Note that "correlated painting" tends not to produce any cross-plane correlation in the beam. See \cite{Wang1999, Hotchi2020}.} The beam was accumulated over 500 injected turns; it was extracted and measured every 50 turns during accumulation. The reconstructed emittances throughout accumulation are shown in Fig.~\ref{fig:7}a.
%
\begin{figure}[b!]
    \centering
    \includegraphics[width=0.5\columnwidth]{fig6.pdf}
    \caption{Measured emittances vs. number of injected turns in the SNS ring. (a) correlated (production) painting at 1 GeV; (b) attempted elliptical painting at 1 GeV; (c)-(d) elliptical painting at 0.8 GeV. In (d), the painted beam size was increased and the beam intensity was decreased relative to (c).}
    \label{fig:7}
\end{figure}
%
As expected, the beam did not develop any significant cross-plane correlation. The beam Twiss parameters were essentially matched at the nominal optics at the RTBT entrance, so bias in the measurement was not a concern.

>>>>

%% For including figures, graphicx.sty has been loaded in
%% elsarticle.cls. If you prefer to use the old commands
%% please give \usepackage{epsfig}
%% The amssymb package provides various useful mathematical symbols
% \usepackage[margin=1in]{geometry}
\usepackage{amssymb}
%% The amsthm package provides extended theorem environments
%% \usepackage{amsthm}
\usepackage{subcaption} % subfigures
\usepackage{gensymb} % \degree symbol
\usepackage{bm}
\usepackage{amsmath}
\usepackage{dblfloatfix} % Put figures at bottom of page
\usepackage{booktabs} % For tables
\usepackage{lineno} % Line numbers
% \linenumbers

@@

\title{Four-dimensional emittance measurement at the Spallation Neutron Source}

@@

\begin{abstract}
A coasting hadron beam with an elliptical transverse profile, uniform charge density, and small transverse four-dimensional (4D) emittance could improve accelerator performance in several contexts. A phase space painting method to generate such a distribution is being tested in the Spallation Neutron Source (SNS) accumulator ring. A critical component of these efforts is to measure the 4D emittance of the beam during accumulation. The 4D emittance can be reconstructed from measured beam profiles in two ways: in the multi-optics method, the optics between a reconstruction and measurement location are varied; in the fixed-optics method, multiple measurement locations are used without modifying the optics. The fixed-optics method is faster but can lead to a large uncertainty in the reconstructed 4D emittance. In this paper, we implement a variant of the multi-optics method using the four available wire-scanners near the SNS target. We also modify the wire-scanner region to reduce the uncertainty of the fixed-optics method. We then demonstrate the usefulness of the fixed-optics method by reconstructing the 4D emittance evolution during accumulation in the SNS ring.
\end{abstract}

@@

The covariance matrix of the transverse phase space distribution of the beam, and hence the root-mean-square emittances, can be reconstructed from 1D projections of the distribution onto a horizontal, vertical, and diagonal axis in the transverse plane \cite{book:Minty2003}.\footnote{
    Reconstruction of the transverse phase space distribution is possible using 2D projections of the distribution on a screen \cite{Hock2013, Wang2019, Wolski2020}. This may be possible using the SNS target imaging system \cite{Blokland2010} but is not pursued here.
}At least four measurements are required, each with a different linear transfer matrix connecting the phase space coordinates at the measurement location to the reconstruction location. The transfer matrix can be controlled by varying the machine optics (multi-optics method), the measurement location (fixed-optics method), or any combination of the two \cite{Prat2014}. The SNS has four wire-scanners available along the transfer line connecting the ring to the spallation target. Thus, the fixed-optics method is possible and preferred in our case due to its speed; however, previous studies have shown that the method can be sensitive to errors \cite{Raimondi1993, Woodley2000, Agapov2007, Faus-Golfe2016}.

The purpose of this work is to optimize the 4D emittance measurement in the SNS. The general reconstruction method is reviewed in Section \ref{sec:Four-dimensional emittance measurement}. The first part of Section \ref{sec:Implementation in the SNS} details the available diagnostics in the SNS; the second part of Section \ref{sec:Implementation in the SNS} describes our implementation of a variant of the multi-optics method in the SNS; the third part of Section \ref{sec:Implementation in the SNS} describes our implementation of the fixed-optics method in the SNS, focusing on reducing the sensitivity to errors. Finally, the usefulness of the fixed-optics method is demonstrated in Section \ref{sec:Example application} by measuring the 4D emittance evolution during accumulation in the SNS ring.

@@

The SNS produces short neutron pulses by colliding a proton beam with a liquid mercury target. The beam starts as a continuous wave of H$^-$ ions which is chopped and then bunched in a 402.5 MHz radio-frequency quadrupole (RFQ), forming microsecond-long minipulses. Each minipulse is accelerated in a linac to a kinetic energy of 1 GeV, stripped of its electrons using a carbon foil, and injected into a ring. Approximately one thousand minipulses are accumulated in the ring over $10^{-3}$ seconds. Finally, the proton beam is extracted and guided through the ring-target beam transport line (RTBT) to the target. This process repeats sixty times per second, resulting in an average beam power of 1.4 megawatts. A comprehensive description of the SNS is given in \cite{Henderson2014}; several design parameters of the SNS ring and RTBT are listed in Table~\ref{tab:SNS}; an overhead view of the ring and RTBT is shown in Fig.~\ref{fig:1}.
%
\begin{figure*}[t!]
    \centering
    \begin{subfigure}{0.35\textwidth}
        % \raisebox{20pt}{\includegraphics[width=\textwidth]{fig1a.pdf}}
        \includegraphics[width=\textwidth]{fig1a.pdf}
    \end{subfigure}
    \hspace{0.25cm}
    \begin{subfigure}{0.615\textwidth}
        \includegraphics[width=\textwidth]{fig1b.pdf}
    \end{subfigure}
    \caption{Left: The accumulator ring and ring-target-beam-transport (RTBT) sections of the Spallation Neutron Source (SNS). Right: $\beta$ functions (top), phase advances (middle), and focusing element and wire-scanner positions (bottom) in the second half of the RTBT. The red box outlines the wire-scanner region.}
    \label{fig:1}
\end{figure*}
%

%
\begin{table}[!b]
    \centering
    \begin{tabular}{ll}
        \toprule
        \textbf{Quantity} & \textbf{Value} \\
        \midrule
        Ring length & 248 [m] \\
        RTBT length & 151 [m] \\
        Beam kinetic energy & 1 [GeV] \\
        Average beam power & 1.4 [MW] \\
        Repetition rate & 60 [Hz] \\
        Transverse tunes ($\nu_x$, $\nu_y$) & (6.23, 6.20) \\
        Longitudinal tune & $7 \times 10^{-4}$ \\
        \midrule 
    \end{tabular}
    \caption{Several design parameters of the SNS ring and RTBT.}
    \label{tab:SNS}
\end{table}
%

Diagnostics in the SNS ring are limited to beam position and loss monitors. Beam profile measurements are performed in the RTBT using four wire-scanners, labeled WS20, WS21, WS23, and WS24, the locations of which are shown in Fig.~\ref{fig:1}. Each wire-scanner consists of three 100$\mu m$-thick tungsten wires mounted on a fork at forty-five-degree angles relative to each other. The intensity of secondary electron emission from each wire is recorded as the fork moves across the beam, thus measuring the 1D projection of the beam distribution onto the $x$, $y$, and $u$ axes, where $u$ is tilted 45 degrees above the $x$ axis. The second-order moments $\langle{xx}\rangle$, $\langle{yy}\rangle$, and $\langle{uu}\rangle$ are estimated from the profiles by averaging over the wire positions weighted by the signal amplitude. The cross-plane moment $\langle{xy}\rangle$ is thus measured indirectly: if the tilt angle of the $u$ axis is $\phi$, then $u = x\cos\phi + y\sin\phi$ and 
%
\begin{equation}
    \langle{xy}\rangle = \frac{\langle{uu}\rangle - \langle{xx}\rangle \cos^2\phi - \langle{yy}\rangle \sin^2\phi}{2\sin\phi\cos\phi}.
\end{equation}
%
The wire-scanners are run in parallel, taking approximately five minutes to move across the beam and return to their original positions. Their default step size is 3 mm and their dynamic range (the ratio of maximum to minimum measured signal) is approximately $100$.\footnote{The root-mean-square beam sizes may depend on beam tails/halo that cannot be measured by the RTBT wire-scanners. Investigation into the effect of beam halo on emittance measurements in the RTBT is left as future work.} They are run at a beam pulse frequency of 1 Hz.\footnote{Each data point corresponds to a separate beam pulse, so the measurement relies on pulse-to-pulse stability.}

@@

We first implemented a variant of the multi-optics method in the SNS. There is considerably more freedom in the choice of optics used in the 4D reconstruction than in 2D reconstruction. We do not consider the general case of coupled optics in this work since the RTBT optics are uncoupled. Thus, the problem reduces to sampling the space of horizontal/vertical phase advances from the reconstruction location to the measurement location such that the reconstruction error is minimized.\footnote{
    In 2D reconstructions, the phase advance from the reconstruction location to the measurement location is typically varied uniformly within a $180\degree$ range. The error is expected to be minimized when the projection angles are spread uniformly between 0 and $\pi$ (the projection angles are only equal to the phase advances in normalized phase space, but it is often convenient to vary the phase advances instead) \cite{Hock2011}. A clear geometric argument is based on the fact that the measured beam sizes define a set of lines that bound an ellipse in 2D phase space at the reconstruction location, so the reconstruction error will be minimized if these lines are evenly distributed around the ellipse. But this argument is less clear in a 4D reconstruction: the measurements of $\left\{\langle{xx}\rangle, \langle{yy}\rangle, \langle{xy}\rangle\right\}$ define a set of surfaces that bound an ellipsoid in 4D phase space at the reconstruction location, and it is not entirely clear how to "evenly distribute" these surfaces around the ellipsoid.
} We chose to follow Prat and Aiba \cite{Prat2014}: in the first half of the scan, the horizontal phase advance is varied while the vertical phase advance is held fixed; in the second half of the scan, the vertical phase advance is varied while the horizontal phase advance is held fixed. Since the four SNS wire-scanners are already spaced somewhat evenly in phase advance, maximal phase coverage is possible by varying the phase advances at each wire-scanner in a $30\degree$ window; however, control of the phase advances between each wire-scanner is limited due to the shared power supplies of the RTBT quadrupoles. There are two power supplies in the wire-scanner region — one controls \{QH18, QH20, QH22, QH24\} and the other controls \{QV19, QV21, QV23, QV25\} — while the last five quadrupoles are controlled independently. In addition to these constraints, the $\beta$ functions must be kept small in the wire-scanner region and must remain close to their nominal values at the target.

@@

This multi-optics method was tested on a fully accumulated production beam in the SNS using ten measurements (forty profiles), which were collected in one hour. The reconstructed Twiss parameters, emittances, and best-fit ellipses in the $x$-$x'$ and $y$-$y'$ planes (normalized by the reconstructed Twiss parameters) are displayed in Fig.~\ref{fig:2}.
%
\begin{figure}[b!]
    \centering
    \includegraphics[width=\columnwidth]{fig2.pdf}
    % \par\medskip
    \begin{tabular}{lll}
        \small\textbf{Parameter} & \small\textbf{Measured} & \small\textbf{Model} \\
        % \midrule
        \small$\beta_x$ [m/rad] & \small22.06 $\pm$ 0.29 & \small22.00 \\
        \small$\beta_y$ [m/rad] & \small4.01 $\pm$ 0.02 & \small3.81 \\
        \small$\alpha_x$ & \small2.33 $\pm$ 0.04 & \small2.37 \\
        \small$\alpha_y$ & \small-0.49 $\pm$ 0.01 & \small-0.60 \\
        \small$\varepsilon_1$ [mm~mrad] & \small33.02 $\pm$ \small0.05 & - \\
        \small$\varepsilon_2$ [mm~mrad] & \small25.67 $\pm$ \small1.03 & - \\
        \small$\varepsilon_x$ [mm~mrad] & \small32.85 $\pm$ \small0.05 & - \\
        \small$\varepsilon_y$ [mm~mrad] & \small25.87 $\pm$ \small0.12 & - \\
      \end{tabular}
    % \par\medskip
    \caption{Reconstructed beam parameters and graphical output (in normalized phase space) from a multi-optics emittance measurement of a production beam.}
    \label{fig:2}
\end{figure}
%
The uncertainties in the parameters were calculated by propagating the standard deviations of the ten reconstructed moments obtained from the LLSQ estimator \cite{Faus-Golfe2016}. The reconstructed Twiss parameters are close to the model parameters computed from the linear transfer matrices of the ring and RTBT, showing that the beam was matched to the nominal optics when it was extracted from the ring. The intrinsic emittances are almost equal to the apparent emittances, showing that there was very little cross-plane correlation in the beam. This is expected for a production beam.

A comprehensive study of errors in the multi-optics 4D emittance measurement was completed at the SwissFEL Injector Test Facility (SITF) by Prat and Aiba in \cite{Prat2014}. They considered errors in the measured moments, quadrupole field and alignment errors, beam energy errors, beam mismatch at the reconstruction point, and dispersion/chromaticity \cite{Mostacci2012}, concluding that the method remained accurate and reporting $< 5\%$ uncertainty in the intrinsic emittances. We initially performed similar studies using envelope tracking to estimate the reconstruction errors in the RTBT \cite{Hoover2021-IPAC}. We considered uncertainty in the beam moments ($\pm 5\%$) quadrupole tilt angles ($\pm\%$ 1 mrad), quadrupole field strengths ($\pm$ 1\%), beam kinetic energy ($\pm$ 3 MeV for a 1 GeV beam), and beam Twiss parameters at the reconstruction point ($\pm$ 5\%), with all other effects turned off. The emittances were successfully reconstructed with systematic and random errors of a few percent. Most of these uncertainties had only a small effect on the reconstructed intrinsic emittances, with the largest coming from beam mismatch and errors in the measured beam size. Therefore, we only consider these sources of uncertainty going forward.\footnote{Space charge forces, which can render the method invalid for high-perveance beams \cite{Anderson2002}, can be neglected since the space charge tune shift in the ring is around 3\% and the distance from the reconstruction location to the measurement location is much less than the length of the ring.} From repeated wire-scanner measurements with the same machine setup (see Fig.~\ref{fig:3}), we estimated the error in the measured moments to be only 3\%. This 3\% noise will be assumed throughout the rest of the analysis.
%
\begin{figure}
    \centering
    \includegraphics[width=0.5\textwidth]{profiles.pdf}
    \caption{Measured profiles of a production beam at WS20 in the RTBT.}
    \label{fig:3}
\end{figure}
%

@@

The downside of the multi-optics method is its long execution time, for the following reasons. First, we are not only interested in the beam emittances at a single time but are also interested in the emittance evolution throughout accumulation, which conveys valuable information about the beam dynamics and allows qualitative comparison with computer simulation. Second, it would be beneficial to quickly evaluate various machine states: although computer simulation can be used as a guide in choosing the experimental parameters, online tuning is expected to be necessary since small changes to the accelerator may have large effects on the painted phase space distribution. Third, the time reserved for accelerator physics experiments at the SNS is limited, and our initial experiments will require substantial setup time due to modification of the beam energy. Thus, the fixed-optics method is preferred. A modest reduction in accuracy for the increase in speed is warranted since weak cross-plane correlations ($\varepsilon_1 \varepsilon_2 \approx \varepsilon_x \varepsilon_y$) are uninteresting for our purposes and do not need to be resolved.

@@

To investigate the sensitivity of the fixed-optics method, we generated a covariance matrix with $\varepsilon_1 = \varepsilon_x$ = 32 mm~mrad and $\varepsilon_2 = \varepsilon_y$ = 25 mm~mrad, close to the measured values, and with Twiss parameters matched to the lattice. This covariance matrix was tracked to the wire-scanners using the transfer matrices from the fifth step in the scan of Fig.~\ref{fig:2}, and the reconstruction was performed many times with 3\% random noise added to the $\langle{xx}\rangle$, $\langle{yy}\rangle$, and $\langle{uu}\rangle$ moments. (We ignored any uncertainty in the transfer matrix elements, which were expected to be small.) This resulted in a large fraction of failed trials, but some successful trials. Fig.~\ref{fig:4} shows the emittances in the successful trials.
%
\begin{figure}[!b]
    \centering
    \includegraphics[width=\columnwidth]{fig3.pdf}
    \caption{Simulated emittance measurement using the nominal optics in the RTBT.}
    \label{fig:4}
\end{figure}
%
Unlike the apparent emittances, the intrinsic emittances are strongly correlated and are not centered on the correct values. Our primary concern is to reduce the latter effect, which we refer to as bias.

@@

Although most of the points are clustered near the original bias and standard deviation of 7\% and 5\%, respectively, the bias increases to nearly 15\% in some cases, which may make it difficult to resolve weak cross-plane correlation; however, the measurement should still resolve strong cross-plane correlation. This is demonstrated in the rest of the plots in Fig.~\ref{fig:6}a, in which the entire process is repeated with $\varepsilon_1 / \varepsilon_2 > 1$. The bias in the reconstruction quickly decreases — the emittances are clustered around their true values. We conclude that with modifications to the RTBT optics, the fixed-optics method should be sufficient for faster (albeit less accurate) four-dimensional emittance measurements in the SNS.

@@

It is first necessary to describe our approach to assigning uncertainty, in the form of error bars, to the emittances reconstructed by the fixed optics method. Propagating the uncertainties in the reconstructed moments, as obtained from the LLSQ estimator, is not an option since no least-squares fit is performed — the cross-plane moments are exactly determined by the four measured profiles. Since the wire-scanners take over five minutes to measure the beam profiles, it is impossible to compute the mean and standard deviation over many trials. We propose to use a similar approach as in the previous section: repeatedly perturb the measured moments in a 3\% range, each time reconstructing the covariance matrix and computing the emittances; compute the mean and standard deviation of the emittances over all trials; plot an error bar centered at the mean, with a width equal to twice the standard deviation.

We first display the measured emittance evolution of a 1 GeV beam produced by the standard SNS painting method.\footnote{The standard painting method is often referred to as "correlated painting". The distance between the injected and circulating beam is varied as the square root of time while the angle between the beams is fixed at zero. Note that ``correlated painting" tends not to produce any cross-plane correlation in the beam. See \cite{Wang1999, Hotchi2020}.} The beam was accumulated over 500 injected turns; it was extracted and measured every 50 turns during accumulation. The reconstructed emittances throughout accumulation are shown in Fig.~\ref{fig:7}a.
%
\begin{figure}[t!]
    \centering
    \includegraphics[width=\columnwidth]{fig6.pdf}
    \caption{Measured emittances vs. number of injected turns in the SNS ring. (a) correlated (production) painting at 1 GeV; (b) attempted elliptical painting at 1 GeV; (c)-(d) elliptical painting at 0.8 GeV. In (d), the painted beam size was increased and the beam intensity was decreased relative to (c).}
    \label{fig:7}
\end{figure}
%
As expected, the beam did not develop any significant cross-plane correlation. The beam Twiss parameters were essentially matched at the nominal optics at the RTBT entrance, so bias in the measurement was not a concern.
